---
title: "Internet Usage Among Latin American People"
author:
    - name: "Eduard Vargas"
      affiliation: "**Monash Student ID: 34322450**"
date: "2024-06-01"
output: 
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


#### Section 1: Loading the libraries
```{r}
# Loading the libraries
library(dplyr)
library(tidyr)
```


#### Section 2: Research Question

Given that technological advancements are often first adopted in developed countries with the necessary resources and infrastructure, this is not typically the case for Latin American countries. What has been the growth in internet usage among people in Latin America from 1990 to the present, and how does it compare to that of a first-world country?

#### Data set introduction:


The data set extracted form [Our World in Data](https://ourworldindata.org/internet).
This is an organisation that focuses on making research and data on major global issues like poverty, disease, hunger, climate change, war, existential risks, and inequality accessible and understandable to promote progress.

On this short data set, the variables contained are the following:

```{r reading the data}
raw_data <- read.csv("number-of-internet-users.csv",header = TRUE)
```

```{r Arranging the names for presentation}
raw_data <- raw_data %>%
        rename(Country = Entity,
               Country_Code = Code,
               Number_of_Internet_Users = Number.of.Internet.users)

```



* **Variables:**








## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
