---
title: "Internet Usage Among Latin American People"
author:
    - name: "Eduard Vargas"
      affiliation: "**Monash Student ID: 34322450**"
date: "2024-06-01"
output: 
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
    number_sections: false
---

```{r setup, include=FALSE}
# Setting up Global Instructions:
knitr::opts_chunk$set(echo = TRUE, include = TRUE, message = FALSE, warning = FALSE)
```


#### Section 1: Loading the libraries
```{r}
# Loading the libraries
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
```


#### Section 2: Research Question

Given that technological advancements are often first adopted in developed countries with the necessary resources and infrastructure, this is not typically the case for Latin American countries. What has been the growth in internet usage among people in Latin America from 1990 to the present, and how does it compare to that of a first-world country?

#### Data set introduction:


The data set extracted form [Our World in Data](https://ourworldindata.org/internet).
This is an organisation that focuses on making research and data on major global issues like poverty, disease, hunger, climate change, war, existential risks, and inequality accessible and understandable to promote progress.

On this short data set, the variables contained are the following:

```{r reading the data}
raw_data <- read.csv("number-of-internet-users.csv",header = TRUE)
```

```{r Arranging the names for presentation, echo=TRUE, include=FALSE}
raw_data <- raw_data %>%
        rename(Country = Entity,
               Country_Code = Code,
               Number_of_Internet_Users = Number.of.Internet.users)

```

```{r}
# Creating a new table with the count of distinct countries
countries_count <- raw_data %>%
  summarise(Count = n_distinct(Country)) %>%
  mutate(Countries = "Number of Countries") %>%
  select(Countries, Count)
```



**Variables:**

The following are the variables contained in the Data Set:
```{r Display Variables}
# Displaying the variables:

print(colnames(raw_data))
```



* **Country:** 
Refers to the 224 distinct Countries contained in the original Data Set.
```{r counting the distinc countries}

# Creating a new table with the count of distinct countries
new_table <- raw_data %>%
  summarise(Count = n_distinct(Country)) %>%
  mutate(Row = "Number of Countries") %>%
  select(Row, Count)

# Displaying the table
kable(new_table, col.names = c("Description", "Count"), caption = "Summary of Data")

```



* **Country_Code:** 

Refers to the specific code allocated for each country

* **Year:** 

Refers to the Years contained.This data contained year 1990 to 2020

```{r Min and Max Years}
# Calculate the minimum and maximum years
min_year <- min(raw_data$Year, na.rm = TRUE)
max_year <- max(raw_data$Year, na.rm = TRUE)

# Create a data frame with these values
min_max_years <- data.frame(
  `Min and Max Years` = c("Minimum Year", "Maximum Year"),
  Year = c(min_year, max_year)
)

# Display the data frame:
kable(min_max_years, col.names = c("Description", "Year"), caption = "Minimum and Maximum Years") %>%
  kable_styling(position = "center")
```

* **Number_of_Internet_Users:** 

Refers to the number of people using (having access) to internet for each year according to the statistics in the data set.



### Section 4: Data set description

The dataset raw_data provides insights into internet usage across different countries. It comprises 6,406 entries, with each entry detailing a specific country, its corresponding code, the year, and the number of internet users. The dataset spans from 1990 to 2020. The number of internet users ranges widely, from 0 to an astounding 4.7 billion, reflecting the exponential growth of internet adoption over the years. The median year in the dataset is 2005, with key statistical quartiles at 1997 and 2012, highlighting significant periods in the dataset's timeline.


```{r Inserting summary Image, pressure, echo=FALSE}
knitr::include_graphics("summary(raw_data).png")
```

**Data types:**

The Raw data contains the following formats:

```{r data formats, fig.align='center'}
str(raw_data)

```


### Section 5: Data summary


In order to resolve the question, I will group the Latin American Countries, and the Northern ones. Subsequently, I will take the mean for each year between these 2 groups and I will compare by year. This will provide an general glimpse on the difference in growth and behavior throughout these time.

```{r Grouping Latin American Countries}

```



